---
title: "Project #1"
author: "Luis R Esteves"
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
library(jsonlite)
library(tidyr)
library(tidyverse)
library(knitr)
library(dplyr)
library(ggplot2)
library(DT)
library(XML)
```

## **Introduction to XML**


### **What are XML files?**

"XML is a markup language that was created by the World Wide WEb Consortium (W3C) to define a syntax for encoding document that both humans and machines could read." In this sense XMl can be considered similar to JSON, since both languages can be interpreted pretty easily by humans. It's very similar to HTML in terms of its stylings. Unlike HTML, XML is a bit more pliable and the use can even create their own markup symbols to describe content. Where HTML is a language that focuses on the presentation of content, XML would be dedicated to data-description for data storage. 


```{r read_in, include=TRUE, message=FALSE}

x <-xmlParse("rows.xml")
rootNode <- xmlRoot(x)
data <- xmlSApply(rootNode, function(x) xmlSApply(x, xmlValue))
data2 <- data.frame(t(data), row.names=NULL)
```